---
title: "??? Introduction to data visualizations in R ???"
author: "Akanksha Garlapati"
date: "`r Sys.Date()`"
output:
  word_document: 
    toc: true
    fig_caption: true
    keep_md: true
    fig_width: 7
    fig_height: 5.5
  pdf_document: default
  html_document: 
    toc: true
    fig_caption: true
    number_sections: true
---

# Introduction to data visualizations in R

## Loading packages and datasets

```{r}
library (tidyverse)
library (modeldata)

library(readxl)
my_dataset <- read_excel("Animals_and_share _of_US_adults.xlsx")
View(my_dataset)
```

## Using scatter plot

We specify dataset to be plotted in ggplot, and the variables and other aesthetics in aes

```{r}
ggplot(my_dataset, aes(x=Male,
    y=Female))
```

We the add the specification for the type of plot with geom\_\*

```{r}
ggplot (my_dataset, aes(x=Male,
    y = Female ))+
    geom_point()
```

We can then add other things like labels

```{r}
ggplot (my_dataset, aes(x = Male,
    y = Female ))+
    geom_point()+
    labs(x="Male",
        y="Female ",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
```

### Specifying other aesthetics

Add color with a variable

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color=Animal ))+
    geom_point()+
    labs(x="Male",
        y="Female",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
```

Label color labels

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point()+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
```

Add color scaling

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point()+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
        #  not found: # soace_color_brewer(palette="Dark2")
```

### Modifying basic properties

Change plot color

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point(color="Red")+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
        # soace_color_brewer(palette="Dark2")
```

change plot size

```{r}
ggplot (my_dataset, aes(x=Male,y=Female,color =Animal))+
    geom_point(size=2)+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
        # soace_color_brewer(palette="Dark2")
```

Adding another plot layer

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point()+
    geom_smooth()+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
    # soace_color_brewer(palette="Dark2")
```

Add geom_smooth method

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point()+
    geom_smooth(method="lm")+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
    # soace_color_brewer(palette="Dark2")
```

Remove geom_smooth error approximation

```{r}
ggplot (my_dataset, aes(x=Male,
    y=Female,
    color =Animal))+
    geom_point()+
    geom_smooth(method="lm",
        se=FALSE)+
    labs(x="Male",
        y="Female",
        color="Animal",
        title="Cricket Chips",
        caption= "Source: Mcdonald (2009)")
    # soace_color_brewer(palette="Dark2")
```

## Other plots

### Histogram

```{r}
ggplot (my_dataset, aes(x=Male,
    color =Animal))+
    geom_histogram()
    
```

Modifying Histogram bins

```{r}
ggplot (my_dataset, aes(x=Male))+
    geom_histogram(bins=15)
    
```

### Frequency polygon

```{r}
ggplot (my_dataset, aes(x=Male))+
    geom_freqpoly(bins=15)
    
```

### Bars

```{r}
ggplot (my_dataset, aes(x=Animal))+
    geom_bar(color="black",
        fill="lightblue")
    
```

Coloring with classes

```{r}
ggplot (my_dataset, aes(x=Animal,
    fill=Animal))+
    geom_bar()
    # soace_fill_brewer(palette="Dark2")
    
```

Remove legend

```{r}
ggplot (my_dataset, aes(x=Animal,
    fill=Animal))+
    geom_bar()
    # soace_fill_brewer(palette="Dark2")
    
```

### Box plot

```{r}
ggplot (my_dataset, aes(x=Animal,y=Female))+
    geom_boxplot()
    # soace_fill_brewer(palette="Dark2")
    
```

## Faceting

```{r}
ggplot (my_dataset, aes(x=Male))+
    geom_histogram(bins=15)+
    facet_wrap(~Animal)
    # soace_fill_brewer(palette="Dark2")
    
```
